Deep Learning

### 2015-04-24 deepreview.tex
more changes to respond to comments from Mikolov and Ranzato

\documentclass[10pts]{article}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{hyperref}

@@

\newcommand{\deri}[2]{\frac{\partial  #1}{\partial #2}}

\parskip 0.085in
\parindent 0.0in
\setcounter{secnumdepth}{1}
\title{Deep Learning}

@@

Deep Learning allows machines that are composed of multiple processing 
layers to learn representions of data with multiple levels of abstraction.
These methods have dramatically improved the state-of-the-art
in speech recognition, object recognition, object detection,
and many other domains such as drug discovery and genomics. Deep
learning discovers intricate structure in large datasets by using the
{\em backpropagation algorithm} to indicate how the machine should
change the weights that are used to compute the representation in each layer from
the representation in the previous layer.  Deep {\em convolutional nets}
have brought about dramatic improvements in processing
images, video, speech and audio, while {\em recurrent
  nets} have shone on sequential data such as text and
speech.
%Convolutional nets have quickly become the leading technology for image
%recognition and speech recognition, while recurrent nets are
%competitive in several natural language processing tasks, including
%language translation. 
%Combinations of recurrent and convolutional nets
%can generate good textual captions of images.

@@

% what is machine learning
Machine learning technology (ML) powers many aspects of modern
society: from web search to content filtering on social networks 
to recommendations on e-commerce websites, and it is increasingly
present in consumer products such as cameras and smartphones. ML
systems are used to identify objects in images, transcribe speech into
text, match news items, posts, or products with users' interests, and
select relevant results of search. Increasingly, these applications
make use of a new class of techniques called {\em deep learning}.

@@

% deep learning: learning representations
Deep learning is a set of methods that allows a machine to be fed with
raw data and to automatically discover the representations needed for
detection or classification. The name comes from the fact that deep learning machines are
multilayer stacks of trainable modules, each of which turns the
representation of the data on its input into a more abstract and
higher-level representation on its output. The output representation
amplifies aspects of the input that are important for discrimination
and suppresses irrelevant variations. An image, for example, comes in the form of an array of
pixel values and the learned features in the first layer of representation
typically represent the presence
or absence of edges at  particular orientations and locations in
the image. The second layer typically detects motifs by spotting particular
arrangements of edges, regardless of small variations in the edge positions
The third layer  may assemble motifs into larger
combinations that correspond to parts of familiar
objects, and subsequent layers would detect objects as combinations of
these parts. The key aspect of deep learning is that these layers of
features are not designed by human engineers: They are learned from
data using a general purpose learning procedure.

% preview of deep learning results
Deep learning is making major advances in solving problems that have
resisted the best attempts of the Artificial Intelligence community
for many years. It has turned out to be very good at discovering
intricate structures in high-dimensional data and is therefore
applicable to many domains of science, business, and government. In
addition to beating records in image
recognition~\citep{Krizhevsky-2012-small,farabet-pami-13,tompson-nips-14,szegedy-2014}
and speech
recognition~\citep{Hinton-et-al-2012,Sainath-et-al-ICASSP2013}, it has
beaten other ML techniques at predicting the activity of potential
drug molecules~\citep{Ma-et-al-2015}, detecting new fundamental
particles~\citep{Ciodaro-et-al-2012,Melis-Higgs-boson-competition-2014},
% YB thinks that Ciodaro-et-al might be {YANNparticles} 
% classifying galaxies, ~\citep{YANNgalaxies}, % YB while waiting for bib entry, 
reconstructing brain circuits~\citep{helmstaedter-nature-2013}, and
predicting the effects of mutations in non-coding DNA on gene
expression and disease~\citep{Keung-et-al-2014,Xiong-et-al-Frey-science2015}.
Perhaps more surprisingly, deep learning has produced extremely
promising results for a variety of tasks in natural language
understanding~\citep{collobert:2011b}, particularly topic
classification, %~\citep{topic-classification},  % YB while waiting for bib entry
sentiment
analysis~\citep{Glorot+al-ICML-2011-small}, 
question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

The backpropagation procedure to compute the gradient of an objective
function with respect to the weights of a multilayer stack of modules
is nothing more than a practical application of the chain rule for
derivatives. The key insight is that the derivative (or gradient) of
the objective with respect to the {\it input} of a module can be
computed by working backwards from the gradient with respect to the
{\em output} of that module (or the input of the subsequent
module), as illustrated in Figure~\ref{fig:backprop-box}.
% Basically, take a module $M$ inside the network, whose input
%is $x$, weight vector is $w$, and output is $y = M(x,w)$. If one knows
%the gradient $\deri{E}{y}$ of the objective function $E$ with respect
%to the module's output $y$, one can blindly apply chain rule to
%compute the gradient with respect to the input:
%\[
%  \deri{E}{x} = \deri{y}{x}\deri{E}{y}
%\]
%This equation works even when $x$ and $y$ are vectors, in which case
%$\deri{E}{x}$ and $\deri{E}{y}$ are gradient vectors, and
%$\deri{y}{x}$ is a {\em Jacobian matrix} of $M$ where element $(i,j)$
%indicates by how much output $j$ would increase proportionally if
%input $i$ were increased by a tiny amount $\Delta y_j =
%\left(\deri{y}{x}\right)_{ij} \Delta x_i$. 
The backpropagation equation can be applied repeatedly to propagate
gradients through all modules, starting from the output at the top 
all the way to the bottom. Once these
gradients have been computed, it is straightforward to compute the gradients 
with respect to the weights of each module. 
%Figure~\ref{fig:backprop-box})
%shows how this works out when the network is composed of layers of
%weight matrices interspersed with scalar non-linearities.
%% the notation in the figure should be updated.

@@

ConvNets exploit the property that many natural
signals are {\em compositional hierarchies}. In images, local
combinations of edges form motifs, motifs assemble into parts, and
parts form objects.  Similar hierarchies
exists in speech and text from sounds to phones, phonemes, syllables,
words, and sentences. The pooling allows representations to vary very little
when elements in the previous layer vary in position and appearance.

@@

Despite these successes, ConvNets were largely forsaken by the
mainstream computer vision and machine learning communities until the
ImageNet competition in 2012.  When deep convolutional networks were
applied to a dataset of about a million images from the web that
contained a thousand different classes they achieved spectacular
results, almost halving the error rates of the best competing
approaches~\citep{Krizhevsky-2012-small}.  This success came from
efficient use of GPUs, ReLUs, and a new regularization technique
called ``dropout'' ~\citep{Srivastava14} which prevented overfitting
in the fully connected layers.  This success has brought about a
revolution in computer vision, and ConvNets are now the dominant
approach for almost all recognition and detection
tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,Taigman-et-al-CVPR2014,simonyan-arxiv-2014,szegedy-2014,tompson-cvpr-15}
and approach human performance on some tasks. A recent stunning
demonstration combines ConvNets and neural net modules for the
generation of image captions, as shown in
Figure~\ref{fig:caption-generation}.

@@

@@

creates a different pattern of activations, or word-vector, illustrated

@@

Vietnam and Iraq. 

@@

%representations allow a lot of parallel computation at the feature
%level so that complex tasks can be solved in a very small number of
%sequential steps.
Deep learning theory shows that deep nets have two different
exponential advantages over classical learning algorithms that do not
use distributed
representations~\citep{Bengio-localfailure-NIPS-2006-small}. Both of
these advantages arise from the power of {\em composition} and depend
on the underlying data generating distribution having an appropriate
componential structure~\citep{Bengio-Courville-Vincent-TPAMI2013}.
First, learning distributed representations enables generalization to
new combinations of the learned features (e.g. $2^n$ combinations,
with $n$ features) ~\citep{Bengio-2009-book,Montufar+Morton-2014}.
Second composing layers in a deep net brings the potential for another
exponential advantage~\citep{Montufar-et-al-NIPS2014} (exponential in
the depth).

@@

Over the last year, several authors have made different proposals to
augment RNNs with a memory module. 
%% This includes simplified
%% version of the LSTM
%% circuit~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014}.
%% The results suggest that not all the elements of the LSTM architecture
%% are required.
Proposals include the ``Neural Turing Machine'' in which the
network is augmented by a ``tape-like'' memory that the RNN
can chose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
``Memory Networks'' in which a regular network is augmented by a
kind of associative memory~\citep{weston-memorynet-2014}. Memory
networks have yielded excellent performance on standard
question-answering benchmarks. The memory is used to remember the
story about which the network is later asked to answer questions.

@@

\bibliography{strings,ml,aigaion,convnetyann,lecun}
\bibliographystyle{natbib}

>>>>

\documentclass[10pts]{article}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
%\usepackage[numbers]{natbib}
\usepackage{hyperref}

@@

\newcommand{\deri}[2]{\frac{\partial  #1}{\partial #2}}
\newcommand{\citep}[1]{\cite{#1}}
\newcommand{\citet}[1]{\cite{#1}}
\parskip 0.085in
\parindent 0.0in
\setcounter{secnumdepth}{1}
\title{Deep Learning}

@@

Deep Learning allows computational models that are composed of multiple processing 
layers to learn representions of data with multiple levels of abstraction.
These methods have dramatically improved the state-of-the-art
in speech recognition, visual object recognition, object detection,
and many other domains such as drug discovery and genomics. Deep
learning discovers intricate structure in large datasets by using the
{\em backpropagation algorithm} to indicate how a machine should
change its internal parameters that are used to compute the representation in each layer from
the representation in the previous layer.  Deep {\em convolutional nets}
have brought about dramatic improvements in processing
images, video, speech and audio, while {\em recurrent
  nets} have shone on sequential data such as text and
speech.
%Convolutional nets have quickly become the leading technology for image
%recognition and speech recognition, while recurrent nets are
%competitive in several natural language processing tasks, including
%language translation. 
%Combinations of recurrent and convolutional nets
%can generate good textual captions of images.

@@

% what is machine learning
Machine learning technology (ML) powers many aspects of modern
society: from web search to content filtering on social networks 
to recommendations on e-commerce websites, and it is increasingly
present in consumer products such as cameras and smartphones. ML
systems are used to identify objects in images, transcribe speech into
text, match news items, posts, or products with users' interests, and
select relevant results of search. Increasingly, these applications
make use of a class of techniques called {\em deep learning}.

@@

% deep learning: learning representations
Representation learning is a set of methods that allows a machine to be fed with
raw data and to automatically discover the representations needed for
detection or classification. Deep learning methods are representation
learning methods with multiple levels of representation, obtained
by composing simple but non-linear modules that each transform the representation at one level
(starting with the raw input) into a representation at a higher, slightly more 
abstract level. With the composition of enough such transformations, very complex functions
can be learned. For classification tasks, the output representation
amplifies aspects of the input that are important for discrimination
and suppresses irrelevant variations. An image, for example, comes in the form of an array of
pixel values and the learned features in the first layer of representation
typically represent the presence
or absence of edges at  particular orientations and locations in
the image. The second layer typically detects motifs by spotting particular
arrangements of edges, regardless of small variations in the edge positions
The third layer  may assemble motifs into larger
combinations that correspond to parts of familiar
objects, and subsequent layers would detect objects as combinations of
these parts. The key aspect of deep learning is that these layers of
features are not designed by human engineers: They are learned from
data using a general purpose learning procedure.

% preview of deep learning results
Deep learning is making major advances in solving problems that have
resisted the best attempts of the Artificial Intelligence community
for many years. It has turned out to be very good at discovering
intricate structures in high-dimensional data and is therefore
applicable to many domains of science, business, and government. In
addition to beating records in image
recognition~\citep{Krizhevsky-2012-small,farabet-pami-13,tompson-nips-14,szegedy-2014}
and speech
recognition~\citep{Mikolov-ASRU-2011,Hinton-et-al-2012,Sainath-et-al-ICASSP2013}, it has
beaten other ML techniques at predicting the activity of potential
drug molecules~\citep{Ma-et-al-2015}, detecting new fundamental
particles~\citep{Ciodaro-et-al-2012,Melis-Higgs-boson-competition-2014},
% YB thinks that Ciodaro-et-al might be {YANNparticles} 
% classifying galaxies, ~\citep{YANNgalaxies}, % YB while waiting for bib entry, 
reconstructing brain circuits~\citep{helmstaedter-nature-2013}, and
predicting the effects of mutations in non-coding DNA on gene
expression and disease~\citep{Keung-et-al-2014,Xiong-et-al-Frey-science2015}.
Perhaps more surprisingly, deep learning has produced extremely
promising results for a variety of tasks in natural language
understanding~\citep{collobert:2011b}, particularly topic
classification, %~\citep{topic-classification},  % YB while waiting for bib entry
sentiment
analysis~\citep{Glorot+al-ICML-2011-small}, 
question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

The backpropagation procedure to compute the gradient of an objective
function with respect to the weights of a multilayer stack of modules
is nothing more than a practical application of the chain rule for
derivatives. The key insight is that the derivative (or gradient) of
the objective with respect to the {\it input} of a module can be
computed by working backwards from the gradient with respect to the
{\em output} of that module (or the input of the subsequent
module), as illustrated in Figure~\ref{fig:backprop-box}.
% Basically, take a module $M$ inside the network, whose input
%is $x$, weight vector is $w$, and output is $y = M(x,w)$. If one knows
%the gradient $\deri{E}{y}$ of the objective function $E$ with respect
%to the module's output $y$, one can blindly apply chain rule to
%compute the gradient with respect to the input:
%\[
%  \deri{E}{x} = \deri{y}{x}\deri{E}{y}
%\]
%This equation works even when $x$ and $y$ are vectors, in which case
%$\deri{E}{x}$ and $\deri{E}{y}$ are gradient vectors, and
%$\deri{y}{x}$ is a {\em Jacobian matrix} of $M$ where element $(i,j)$
%indicates by how much output $j$ would increase proportionally if
%input $i$ were increased by a tiny amount $\Delta y_j =
%\left(\deri{y}{x}\right)_{ij} \Delta x_i$. 
The backpropagation equation can be applied repeatedly to propagate
gradients through all modules, starting from the output at the top 
(where the network produces its prediction) all the way to the bottom
(where the external input is fed). Once these
gradients have been computed, it is straightforward to compute the gradients 
with respect to the weights of each module. 
%Figure~\ref{fig:backprop-box})
%shows how this works out when the network is composed of layers of
%weight matrices interspersed with scalar non-linearities.
%% the notation in the figure should be updated.

@@

Deep neural networks exploit the property that many natural
signals are {\em compositional hierarchies}, where higher-level
features are obtained by composing lower-level ones. In images, local
combinations of edges form motifs, motifs assemble into parts, and
parts form objects.  Similar hierarchies
exists in speech and text from sounds to phones, phonemes, syllables,
words, and sentences. The pooling allows representations to vary very little
when elements in the previous layer vary in position and appearance.

@@

Despite these successes, ConvNets were largely forsaken by the
mainstream computer vision and machine learning communities until the
ImageNet competition in 2012.  When deep convolutional networks were
applied to a dataset of about a million images from the web that
contained a thousand different classes they achieved spectacular
results, almost halving the error rates of the best competing
approaches~\citep{Krizhevsky-2012-small}.  This success came from
efficient use of GPUs, ReLUs, a new regularization technique
called ``dropout'' ~\citep{Srivastava14}, and techniques to
generate more training examples by deforming the existing ones.
This success has brought about a
revolution in computer vision, and ConvNets are now the dominant
approach for almost all recognition and detection
tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,Taigman-et-al-CVPR2014,simonyan-arxiv-2014,szegedy-2014,tompson-cvpr-15}
and approach human performance on some tasks. A recent stunning
demonstration combines ConvNets and neural net modules for the
generation of image captions, as shown in
Figure~\ref{fig:caption-generation}.

@@

%representations allow a lot of parallel computation at the feature
%level so that complex tasks can be solved in a very small number of
%sequential steps.
Deep learning theory shows that deep nets have two different
exponential advantages over classical learning algorithms that do not
use distributed
representations~\citep{Bengio-localfailure-NIPS-2006-small}. Both of
these advantages arise from the power of {\em composition} and depend
on the underlying data generating distribution having an appropriate
componential structure~\citep{Bengio-Courville-Vincent-TPAMI2013}.
First, learning distributed representations enables generalization to
new combinations of the values learned features beyond those seen
during training (e.g. $2^n$ combinations
are possible with $n$ binary features) ~\citep{Bengio-2009-book,Montufar+Morton-2014}.
Second composing layers of representation in a deep net brings the potential for another
exponential advantage~\citep{Montufar-et-al-NIPS2014} (exponential in
the depth).

@@

creates a different pattern of activations, or word-vector, a distributed
representation illustrated

@@

Sweden and Norway.

@@

@@

Over the last year, several authors have made different proposals to
augment RNNs with a memory module. 
%% This includes simplified
%% version of the LSTM
%% circuit~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014}.
%% The results suggest that not all the elements of the LSTM architecture
%% are required.
Proposals include the ``Neural Turing Machine'' in which the
network is augmented by a ``tape-like'' memory that the RNN
can choose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
``Memory Networks'' in which a regular network is augmented by a
kind of associative memory~\citep{weston-memorynet-2014}. Memory
networks have yielded excellent performance on standard
question-answering benchmarks. The memory is used to remember the
story about which the network is later asked to answer questions.

@@

\bibliography{strings,ml,aigaion,convnetyann,lecun}
%\bibliographystyle{natbib}
\bibliographystyle{ieeetr}
